### 2020-02-20 docs/papers/osr/body.tex
An initial outline.

\newcommand{ironstar}{Iron{\star}}
\newcommand{bepsilon}{B^{\epsilon}}

Systems verification is picking up momentum.
The promise is that we might someday replace a methodology of
write-test-debug with a comprehensive compile-time confirmation
of correctness. In particular, critical software could remove
``ship'' from the bug-fix development cycle.

The chief obstacle to the practical application of verification for
systems-scale software development is the sheer tedium of involved
in reducing a proof to the level of mechanically verifiability.
Automated verifiers based on SAT solvers have made tremendous progress
grinding through this logical tedium, brinigng systems-scale
verification into the realm of feasibility.

\section{{\ironstar}: an automation-based approach}

The authors have been involved in a few projects~\cite{}
over which we have developed strategies that exploit
SAT-solver automation to improve the scalability of verification.

\section{What makes systems verification challenging?}

\subsection{Stating a system specification}

\subsection{Structuring a proof for maintainability}

\subsection{Proof tedium}

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{How {\ironstar} addresses these challenges}

\subsection{Stating a system specification}\label{sec:specification}

{ironstar} adopts Lamport's TLA+ model for specifying the
allowed behaviors of an evolving system~\cite{specifying-systems}.
TLA+ emphasizes
\begin{itemize}
\item Mathematical abstraction to elide every detail that isn't
  essential to a specification
\item Nondeterminism to capture environmental concurrency,
  nondeterminism, adversarial behavior, and abstract 
\end{itemize}

TLA+ is built on untyped set theory, but we find using a
decidable type system dramatically more convenient,
as it dispatches many tedious proof obligations in the type checker before
reaching the theorem prover.
Dafny's type system offers numeric and logical primitive types,
sequences, sets and maps, user-defined
algebraic (struct/tagged-union) recursive types, and generics.

We define behaviors as TLA+-style state machines:
a struct provides the \textemph{state} type, and
then \textsc{Init} and \textsc{Next} predicates define
the allowed nondeterministic state machine relations.

A state machine is a natural way to define the goalposts for a system.
How should a replicated state machine behave?
Just like a instance of single state machine.
How should a physically decentralized filesystem behave?
Just like a logically-centralized filesystem~\cite{Farsite}.
How should a {\bepsilon}tree store behave?
Just like a simple map.

A state-machine goal avoids a common verification trap of specifying
an implementation's correctness by its invariants. The invariants are
means to an end; a reference state machine is often the most straightforward
way to describe that end.

\subsection{Modeling the environment}

A distributed system consists of several processes running
the program we wish to verify,
a network,
and our assumption that those the processes may only interact through
the network.

A storage system consists of a process running the program
we wish to verify,
a storage device,
our assumption about the asynchronous bus that connects the
device to the process,
and our assumption that the system may sponatenously \texttt{crash},
zeroing the process memory but preserving the device memory.
\footnote{
It may be easier to see how general the model is if you imagine
the storage system as a strangely-shaped distributed system:
one process that runs the program,
another that runs a trusted storage state machine,
a network (the SATA bus) that connects them,
and a failure model for the processes.
The same sorts of nondeterminism that reorders network messages and fails
hosts in a distributed system
can reorder I/O requests and fail the filesystem process.
}

In both cases, the ultimate theorem claims that the
\textit{system} behavior---process(es) and environment---refines
to that of the logically-centralized specification.
What does mechanical verification of the program mean?
It means that, if the program is instantiated as processes in an
environment matching the \textit{environment model},
its behavior will match the \textit{application model}.
That is, one must read the text defining not only the application
but also the environment to know what the verification promises;
in exchange, one needn't read the program implementation, as the
mechanical verifier has done that for you.

\subsection{Structuring a proof for maintainability}

A systems verification problem, ultimately, is simply a really complex
proof goal: an inductive statement of refinement between an implementation
state machine and a specification state machine, each of which is a
big tree of definitions ultimately rooted in some first-order logic or
basic datatypes.

In early interactive theorem provers, the operator stated a goal theorem.
The proof engine offered a bag of tools that transform the proof rules.
The operator repeatedly applied tools until the proof goal has been
transformed into \texttt{true}.
These systems would record the operator's effort as a ``proof script'';
one could replay the script to verify the veracity of the top-level goal.

Proof scripts are brittle; a small change in the system definitions perturb
the proof, and necessary repairs are often non-local~\cite{}.
Chlipala has made remarkable progress in organizing proof scripts into
tactic modules that mimic the structure of the system under
verification. His approach improves the modularity and locality
of maintenance of the proof artifact.
It requires the operator to simultaneously maintain the system
and the tactic library that supports it.

\comment{Diagram here.}

An {\ironstar} proof is a refinement hierarchy of state machines.

In {\ironstar}, the operator works as follows:
First, she defines the state machine (and its data structures) that
specify the application~(\S\ref{sec:specification}).
She uses abstraction to omit as many details as possible while capturing the
desired application behavior.
She also models the finished \texttt{system} as a compound state machine that
models the environment and takes a program as a parameter.
Her goal is to write a program that, when instatiated in the \texttt{system},
causes the system to satisify the application spec.

The program will ultimately be expressed as
imperative code over an immutable heap,
so that the programmer has sufficient control to achieve the
desired performance.
The imperative code is event-driven, so that we can view each
trip through the main loop or invocation of a handler as an
atomic step in a state machine,
and thus can be plugged in as a component into the \texttt{system} definition.

In any system of reasonable size, the semantic gap between the abstract
application definition and the pointer-diddling implementation will be
substantial.
Why is any given line of C++ code correct? There are surely many reasons.
Perhaps the \texttt{->} operator isn't a null pointer dereference because
the pointer was freshly allocated six lines ago; an automatic tool may
be able to confirm such local or simple properties.
But the arguments to the referenced method might be the right ones because
they were the results of another method call earlier in the function.
And the side-effect of the method call might be correct
because it establishes a property that will be used when another I/O
event completes in the future.
Most of these higher-level arguments appear at best in comments,
or perhaps in a distant and outdated English design document,
or they can be inferred from the penumbras and emanations of the
call graph or object hierarchy.
Most likely, they appear nowhere other than an author's mental model.

In a verified system, every such correctness argument must be explicitly
stated and the collection chained together to connect the implementation
code all the way up to the abstract application specification.
In an {\ironstar} verified system, these arguments are organized into
a refinement hierarchy.
In the refinement hierarchy, state machine models act as waypoints
along the route from abstract to concrete. Each refinement proof that
connects a state machine to the abstract one above it captures some
aspect of the system's correctness reasoning. 
The refinement hierarchy is a mechanism to \textit{organize the
proof arguments} into a maintainable structure.

The idea of refinement as a proof strategy is far from new with {\ironstar}.
Note that we do \textit{not} aim to use refinement
as a way to break a proof down into small-enough pieces that a
mechanical verifier can leap the layers completely automatically
(as is done in~\cite{cspec,ironarmada}).
Instead, the refinement layers are meant to be ``human-sized'' concepts:
just as a skilled engineer decomposes a function or organizes a
class hierchy at a scale that's designed to communicate to the next
human reader,
refinement layers in {\ironstar} are designed to organize the proof
artifact into human-brain-sized conceptual chunks.

\comment{Discuss the tree hierarchy in veribetrfs}
Figure~{diagram} shows the richest refinement hierarchy in the
{\veribetrfs} system.
The top refinement explains how queries searching a {\bepsilon}tree
structure give it behavior equivalent to a map,
while abstracting the internal structure of tree nodes as infinite
maps.
The next layer explains how nodes can be implemented in finite space
with pivots; note that this layer disregards the {\bepsilon}tree structure,
since those correctness arguments are already dispatched in the
layer above.
The pivot layer assumes that nodes are abstract datatypes that can
be passed through disk storage, which of course is an abstraction of
the real disk. The next layer shows how marshalling can provide the
same abstract data types while passing byte strings through disk storage.

Each of the layers above are written in with immutable data types
(that is, in a functional, side-effect-free style), abstracting away
the complexity of in-place updates. They also make widespread use of
mathematical integers, wherein $i+1$ is always bigger than $i$.
The bottom layer -- the concrete implementation -- uses machine integers
and exploits in-place updates to achieve excellent performance.
Its refinement proof shows that both transformations are correct,
but again, it disregards any of the reasoning at higher levels of
abstraction.

The developer uses

\subsection{Proof tedium}

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{Managing automation}