### 2020-02-27 docs/papers/osr/body.tex
a draft

\newcommand{ironstar}{Iron{\star}}
\newcommand{bepsilon}{B^{\epsilon}}
\newcommand{veribetrfs}{\texttt{\textit{veri}betrfs}}

For half a century, we have known that a program is constructed by
writing code, testing, and debugging. Program verification promises
an alternative process in which no bug escapes the compiler.

Convincing a mechanical checker that a program is correct requires
resolving a fantastic degree of detail.
The SMART system offered an extension to Paxos~\cite{smart}.
Its authors provided a hand-written proof at a level of detail intended
to satisfy a mechanical checker, although it was never checked.
Just the proof of the SMART algorithm, without an imperative implementation,
is 70 pages of text~\cite{smart-proof}.
If we are to apply verification as a practical engineering tool,
we must rely heavily on automation to retire tedious logic obligations,
focusing humans on just the subtle invariants that reflect the insights
of the system design.

The authors have verified several systems using highly-automated
verification tools. This paper conveys the experience of learning and using
such tools. We argue that strong automation is the path forward for
large-scale systems verification, and we highlight open challenges.

\section{A verified codebase}

The code base of a verified system has three parts:
the specification, the implementation, and the proof.

The \textit{specification} defines the application spec
(what does the program do?) and the environment spec
(what can the program assume about its environment?).

The \textit{implementation} is the source code that will compile
into the final program. It is imperative (has in-place updates) and
uses machine-supported datatypes (uint64) to give the programmer
enough control to achieve great performance.

The \textit{proof} files illuminate the invariants and structure
the arguments that show that the implementation does what the specification
says.

A typical specification is smaller than 1/10^{th} the size of
the implementation it constrains.
In {\ironstar}, the spec files carry a distinguished file extension,
and spec files may only include other spec files.
If the developer reads all of the spec files, she need read no other
files to understand what the verified program does.

An {\ironstar} proof text is around five times
as big as the implementation code.
The effort spent constructing and maintaining it is in lieu of
maintaining a conventional test suite.

\section{A day in the life of a verification engineer}

At the large scale, the team's task is to write a fast implementation,
and show that its behavior matches the spec. This task breaks down into
making smaller arguments; a typical argument might say ``splitting a node
in a tree creates a new tree that represents all the same key-value pairs.''

When using an automation-centric tools, the engineer begins each such
argument the same way: She states the argument, and then offers an empty
proof:
\begin{verbatim} % TODO get an environment that numbers lines.
lemma TODOPlausibleLemmaName(before:Tree, after:Tree, pivot:Key)
  requires TreeInv(a)
  requires SplitNodeRelation(before, after, pivot)
  ensures TreeInv(b)
{
}
\end{verbatim}
Then she asks the verifier to try to verify the lemma.
This is analogous to testing every possible argument assignment,
but of course the mechanism is not actually model checking, since
there are infinite such assignments.

Sometimes, automation
successfully opens the tall tree of definitions under
\texttt{SplitNodeRelation} and \texttt{TreeInv}, and confirms the validity
of the lemma.
More commonly, the verifier cannot convince itself, and it reports
\texttt{TODO real text at line 4}.
Maybe the lemma is invalid: The definitions don't actually
do what the engineer hoped, either because of a deep misunderstanding
or due to a minor off-by-one error.
This is why we're using automation in the first place; it's a failed test
pointing to a flaw in the designer's understanding.
Or perhaps the lemma is valid, but the automation can't see why.
This is by far the more common case; it's the tax we pay to enjoy
exhaustive testing.

However, the engineer's action in either case is the same: she must
elaborate the proof until either the verifier admits its validity
or she discovers an underlying error.

The currency of definitions is are quantified logical statments.
\verb{TreeInv(t)} is a list of conjuncts; one says that the tree
is acyclic:
\begin{verbatim}
predicate Acyclic(t) {
  forall p, i, j :: Path(p, t) && i!=j ==> p[i] != p[i]
}
\end{verbatim}
This says that no path through the tree visits the same node twice.

Suppose the engineer wonders if acyclicity is the thing tripping up the
verifier. So she replaces the lemma proof with:
\begin{verbatim}
{
  assert Acyclic(b);
}
\end{verbatim}
The verifier will respond in one of the following ways:
\begin{itemize}
  \item line 6 assertion failed: Yep, acyclicity is the difficult part,
  \textit{and if we assume it's true, the postcondition verifies}.
  \item line 4 postcondition failed: acyclicity was valid; some other
  conjunct in TreeInv(b) is the problem.
  \item \textit{both} errors: The verified needs help seeing more than one
  conjunct in \texttt{TreeInv}.
  \item neither error: the assertion itself has given the verifier a
  sufficient hint to verify the lemma (see \S\ref{observes}).
  \item \textit{Timeout}. See \S\ref{timeouts}.
\end{itemize}
The first two cases are by far the most common: the engineer has successfully
"bisected" the problem, either implicating or vindicating the \verb{Acyclic}
conjunct.

She applies this bisection technique recursively, just as she would use
printf bisection debugging to back a deterministic runtime test failure into a
corner.
If \verb{Acyclic(b)} is unproven, she drills down:
\begin{verbatim}
{
  forall p, i, j | Path(p, t) && i!=j
    ensures p[i] != p[i]
  {
  }
  assert Acyclic(b);
}
\end{verbatim}
The forall brings new variables \verb{p, i, j} into scope for use in its
proof body which, again, begins as \verb{\{\}}.
The verifier complains about the postcondition on line 7.
It accepts the assert on line 10, predicated on the assumption that the
engineer will ultimately correct the failing forall statement.

Now that \verb{p, i, j} have been introduced, the engineer may drill
further down into definitions to argue why \verb{p[i] != p[i]}.
In this exmaple, she would begin a proof by contradiction:
If there exists a cyclic path \verb{pb} in \verb{b},
then construct a witness path \verb{pa} in tree \verb{a} that is
cyclic, which contradicts the given \verb{Acyclic(a)}.

This iterative, user-guided debugging approach stands in contrast to other
common verification development methodologies.
The engineer isn't confronted with a half-transformed proof goal
with a hundred subterms, forcing her to
think in the verifier's internal representation.
Nor does she dig through a counterexample with a value assigned for each
of hundreds of in-scope variables, only two of which are actually relevant.
Both of those approaches present information to the user that scales up with
the system size, and both ask the user to interpret the system's model.
By contrast, bisection proof debugging lets the engineer develop her own
mental model by probing the verification Oracle in a game of twenty-questions.
The benefit of letting the engineer steer the story and maintain her own
mental model outweighs the fact that she must make a logarithmic number of
inquiries to reach the root cause, especially because each inquiry is answered
rapidly.

Returning to the possible verifier responses: in case 3, the verifier
complains about both the assertion and the postcondition; the engineer
has failed to bisect the problem. It could be as simple as the assertion
being incorrect, or it could be that there is more than one problem
(error in definition, or missing proof) to identify. The engineer proceeds
the same as above, but has more ground to cover before she's done.

Case 4 is a victory: by introducing the assertion, the engineer has
triggered the automation to exploit a quantifier it hadn't considered
automatically, and it has completed the proof.

Case 5, a timeout, is a meta-problem: not only can the verifier not
see the proof, but the automation strategy has broken down, and requires
human intervention to fix. These issues are rare: we bisect-debug all day,
but have to resolve a timeout perhaps once a week.
We discuss automation control and timeout mitigation in Section~\ref{timeout}.

Other than timeout management, note that the nitty-gritty work of verification
engineering feels a lot like printf-debugging a deterministic test failure.
The process is familiar, easy to learn, and it generally requires deep insight
only about the system being debugged, not about the verification tool itself.

TODO the glory of quantified vs recursive definitions.

\subsection{Structuring the proof}

While the low-level mechanics are simple, the verification engineer does
need to adopt new disciplines to organize her proof text for scale and
maintainability. The goal is the same as organizing a big software system
into modules for maintainability, but the techniques for arranging a
proof argument are different.

A single line of an imperative program may be correct for many reasons,
at multiple levels of abstraction, all of which must hold to support the
program's correctness. For example, in this line:
\begin{verbatim}
  node->writeToBlock(lba)
\end{verbatim}
Dereferencing \verb{node} is safe because \verb{node} was allocated earlier
in the method and not freed. The write to address \verb{lba} is safe
because \verb{lba} was returned from a library function managing free space
on the device. The content written, a marshalled representation of \verb{node},
is appropriate because later in the program we'll store a pointer to \verb{lba}
that expects a representation of \verb{node} to appear there.

Common static analysis tools can confirm non-null properties like the first
one. The other arguments are at higher levels of abstraction, interact with
distant lines of code in the program, and may even interect with other
instances of the process, such as before or after a crash event that resets
RAM. In a conventional program, each informal argument might appear in a
comment on the code line, or might be inferrable from the name of
the methods or the methods it calls, or may appear in an Engilsh design
document from two years ago, or may appear nowhere but in an engineer's
memory.

Mechanical verification requires that each such argument be written down.
Trying to cram all of these arguments into the ground-level source code
is overwhelming; the more abstract arguments are weighed down with the
tedium of implementation. In {\ironstar}, we employ state machine refinement
to organize proof text%
\footnote{We contrast this use with other verification efforts that also
use refinement, but in service to the verification system, to divide the
proof up into layers so fine that automation can leap from one to the next
without assistance. In {\ironstar}, refinement layers are introduced at
a granularity that humans find natural, just as developers choose appropriate
granularity boundaries for functional abstraction or object encapsulation.
}
by separating it along the axis of abstraction.

% TODO discuss app specification in more detail somewhere?
The app spec is a TLA+-style state machine.
In {\veribetrfs}, the spec is a key-value dictionary extended with
a \verb{flush} operation and a \verb{crash} event that interleaves
at the whim of the environment. These specs are all written with immutable
data structures and mathematical conveniences (unbounded integers, infinite
sets and sequences) that let us focus on the behavior while ignoring
implementation details.

The first layer of refinement introduces the structure of a {\bepsilon}
tree, while abstracting the nodes as infinite sets mapping every key to
a child node.
Proof text at this layer argues that the tree state machine maintains the
necessary invariants such that the tree acts exactly like a dictionary.

The second layer refines the infinite nodes to ``pivot'' nodes, which
map keys from A---G to one child, H---N to the next, and so on. Proof
text at this layer argues that pivot nodes are a (finite!) shorthand
for infinite maps, and hence that {\bepsilon}{\langle}pivot{\rangle} is
still a dictionary.

The previous layer supposed a disk that accepts the Node abstract data
type as arguments to \verb{write}; the third layer refines the Pivot Node to
have a marshalled representation. Proof text at this layer argues
that \verb{unmarshall(marshall(n) == n}, and thus that, when
a {\bepsilon} tree made of marshallable nodes is connected to a disk
that does I/O on byte strings, the system behaves the same as the prior layer.

Finally, the bottom-most layer replaces immutable datatypes with mutable
ones (to exploit the performance of in-place updates), and replaces
mathematical integers with machine \verb{uint64}s. Proof text at this
layer shows that aliasing is controlled enough that mutable data structures
mimic their immutable models above. It also shows that bounds are respected
so that \verb{uint64}s don't overflow.

Each of the arguments is necessary to stitch the proof together from the bottom
layer implementation to the top level application spec.
The important observation from this tour is that the refinement layers
act as narrow interfaces along the axis of abstraction: arguments about
high-level concepts like {\bepsilon} tree structure may omit details
about marshalling and register bounds; conversely, low-level arguments
may attend only to detailed reasoning, as the abstract properties were
dispensed with elsewhere.

Module design in an imperative program creates conceptual boundaries,
so that an engineer can study and modify part of the program while being
insulated from interactions with most of the other lines of code in the
system.
Likewise, we have found this organization enables us to study and modify
parts of our verified system while insulating our changes from other
parts. Most importantly, it means that we can introduce a new feature
(like sub-node reads) while leaving most of the proof text undisturbed.
Because proof is 80\% of the text of our verified system,
incremental and modular maintenance is as important for the proof as
for the imperaive code.

------------------------------------------------------------------------------
Systems verification is picking up momentum.
The promise is that we might someday replace a methodology of
write-test-debug with a comprehensive compile-time confirmation
of correctness. In particular, critical software could remove
``ship'' from the bug-fix development cycle.

The chief obstacle to the practical application of verification for
systems-scale software development is the sheer tedium of involved
in reducing a proof to the level of mechanically verifiability.
Automated verifiers based on SAT solvers have made tremendous progress
grinding through this logical tedium, brinigng systems-scale
verification into the realm of feasibility.

\section{{\ironstar}: an automation-based approach}

The authors have been involved in a few projects~\cite{}
over which we have developed strategies that exploit
SAT-solver automation to improve the scalability of verification.

\section{What makes systems verification challenging?}

\subsection{Stating a system specification}

\subsection{Structuring a proof for maintainability}

\subsection{Proof tedium}

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{How {\ironstar} addresses these challenges}

\subsection{Stating a system specification}\label{sec:specification}

{ironstar} adopts Lamport's TLA+ model for specifying the
allowed behaviors of an evolving system~\cite{specifying-systems}.
TLA+ emphasizes
\begin{itemize}
\item Mathematical abstraction to elide every detail that isn't
  essential to a specification
\item Nondeterminism to capture environmental concurrency,
  nondeterminism, adversarial behavior, and abstract 
\end{itemize}

TLA+ is built on untyped set theory, but we find using a
decidable type system dramatically more convenient,
as it dispatches many tedious proof obligations in the type checker before
reaching the theorem prover.
Dafny's type system offers numeric and logical primitive types,
sequences, sets and maps, user-defined
algebraic (struct/tagged-union) recursive types, and generics.

We define behaviors as TLA+-style state machines:
a struct provides the \textemph{state} type, and
then \textsc{Init} and \textsc{Next} predicates define
the allowed nondeterministic state machine relations.

A state machine is a natural way to define the goalposts for a system.
How should a replicated state machine behave?
Just like a instance of single state machine.
How should a physically decentralized filesystem behave?
Just like a logically-centralized filesystem~\cite{Farsite}.
How should a {\bepsilon}tree store behave?
Just like a simple map.

A state-machine goal avoids a common verification trap of specifying
an implementation's correctness by its invariants. The invariants are
means to an end; a reference state machine is often the most straightforward
way to describe that end.

\subsection{Modeling the environment}

A distributed system consists of several processes running
the program we wish to verify,
a network,
and our assumption that those the processes may only interact through
the network.

A storage system consists of a process running the program
we wish to verify,
a storage device,
our assumption about the asynchronous bus that connects the
device to the process,
and our assumption that the system may sponatenously \texttt{crash},
zeroing the process memory but preserving the device memory.
\footnote{
It may be easier to see how general the model is if you imagine
the storage system as a strangely-shaped distributed system:
one process that runs the program,
another that runs a trusted storage state machine,
a network (the SATA bus) that connects them,
and a failure model for the processes.
The same sorts of nondeterminism that reorders network messages and fails
hosts in a distributed system
can reorder I/O requests and fail the filesystem process.
}

In both cases, the ultimate theorem claims that the
\textit{system} behavior---process(es) and environment---refines
to that of the logically-centralized specification.
What does mechanical verification of the program mean?
It means that, if the program is instantiated as processes in an
environment matching the \textit{environment model},
its behavior will match the \textit{application model}.
That is, one must read the text defining not only the application
but also the environment to know what the verification promises;
in exchange, one needn't read the program implementation, as the
mechanical verifier has done that for you.

\subsection{Structuring a proof for maintainability}

A systems verification problem, ultimately, is simply a really complex
proof goal: an inductive statement of refinement between an implementation
state machine and a specification state machine, each of which is a
big tree of definitions ultimately rooted in some first-order logic or
basic datatypes.

In early interactive theorem provers, the operator stated a goal theorem.
The proof engine offered a bag of tools that transform the proof rules.
The operator repeatedly applied tools until the proof goal has been
transformed into \texttt{true}.
These systems would record the operator's effort as a ``proof script'';
one could replay the script to verify the veracity of the top-level goal.

Proof scripts are brittle; a small change in the system definitions perturb
the proof, and necessary repairs are often non-local~\cite{}.
Chlipala has made remarkable progress in organizing proof scripts into
tactic modules that mimic the structure of the system under
verification. His approach improves the modularity and locality
of maintenance of the proof artifact.
It requires the operator to simultaneously maintain the system
and the tactic library that supports it.

\comment{Diagram here.}

An {\ironstar} proof is a refinement hierarchy of state machines.

In {\ironstar}, the operator works as follows:
First, she defines the state machine (and its data structures) that
specify the application~(\S\ref{sec:specification}).
She uses abstraction to omit as many details as possible while capturing the
desired application behavior.
She also models the finished \texttt{system} as a compound state machine that
models the environment and takes a program as a parameter.
Her goal is to write a program that, when instatiated in the \texttt{system},
causes the system to satisify the application spec.

The program will ultimately be expressed as
imperative code over an immutable heap,
so that the programmer has sufficient control to achieve the
desired performance.
The imperative code is event-driven, so that we can view each
trip through the main loop or invocation of a handler as an
atomic step in a state machine,
and thus can be plugged in as a component into the \texttt{system} definition.

In any system of reasonable size, the semantic gap between the abstract
application definition and the pointer-diddling implementation will be
substantial.
Why is any given line of C++ code correct? There are surely many reasons.
Perhaps the \texttt{->} operator isn't a null pointer dereference because
the pointer was freshly allocated six lines ago; an automatic tool may
be able to confirm such local or simple properties.
But the arguments to the referenced method might be the right ones because
they were the results of another method call earlier in the function.
And the side-effect of the method call might be correct
because it establishes a property that will be used when another I/O
event completes in the future.
Most of these higher-level arguments appear at best in comments,
or perhaps in a distant and outdated English design document,
or they can be inferred from the penumbras and emanations of the
call graph or object hierarchy.
Most likely, they appear nowhere other than an author's mental model.

In a verified system, every such correctness argument must be explicitly
stated and the collection chained together to connect the implementation
code all the way up to the abstract application specification.
In an {\ironstar} verified system, these arguments are organized into
a refinement hierarchy.
In the refinement hierarchy, state machine models act as waypoints
along the route from abstract to concrete. Each refinement proof that
connects a state machine to the abstract one above it captures some
aspect of the system's correctness reasoning. 
The refinement hierarchy is a mechanism to \textit{organize the
proof arguments} into a maintainable structure.

The idea of refinement as a proof strategy is far from new with {\ironstar}.
Note that we do \textit{not} aim to use refinement
as a way to break a proof down into small-enough pieces that a
mechanical verifier can leap the layers completely automatically
(as is done in~\cite{cspec,ironarmada}).
Instead, the refinement layers are meant to be ``human-sized'' concepts:
just as a skilled engineer decomposes a function or organizes a
class hierchy at a scale that's designed to communicate to the next
human reader,
refinement layers in {\ironstar} are designed to organize the proof
artifact into human-brain-sized conceptual chunks.

\comment{Discuss the tree hierarchy in veribetrfs}
Figure~\ref{betreerefinement} shows the richest refinement hierarchy in the
{\veribetrfs} system.
The top refinement explains how queries searching a {\bepsilon}tree
structure give it behavior equivalent to a map,
while abstracting the internal structure of tree nodes as infinite
maps.
The next layer explains how nodes can be implemented in finite space
with pivots; note that this layer disregards the {\bepsilon}tree structure,
since those correctness arguments are already dispatched in the
layer above.
The pivot layer assumes that nodes are abstract datatypes that can
be passed through disk storage, which of course is an abstraction of
the real disk. The next layer shows how marshalling can provide the
same abstract data types while passing byte strings through disk storage.

Each of the layers above are written in with immutable data types
(that is, in a functional, side-effect-free style), abstracting away
the complexity of in-place updates. They also make widespread use of
mathematical integers, wherein $i+1$ is always bigger than $i$.
The bottom layer -- the concrete implementation -- uses machine integers
and exploits in-place updates to achieve excellent performance.
Its refinement proof shows that both transformations are correct,
but again, it disregards any of the reasoning at higher levels of
abstraction.

The developer creates the intermediate state machines as
``interface boundaries'' that separate concerns.
She organizes the bulk of the proof artifact
by slicing it across the axis of abstraction.

Of course, the proof artifact can also be orthogonally organized
by module.
In {\veribetrfs},
the {\bepsilon}tree refinement stack of Figure~\ref{betreerefinement}
has a (simpler) sibling that reasons about a disk block caching module.

\subsection{Proof tedium}

\subsubsection{A day in the life of a verification engineer.}
What is it like to develop software in this style?

Define a state machine.
- it's immutable datatypes; get comfortable with abstraction.
State a refinement.
- it's a boilerplate induction pattern.
Assert its proof.
- Start with {} -- see how far automation takes you.
- Probably a simple case breakdown. Write an if or case statement!
Diagnose
- bisection debugging! This is an exciting part of the story;
  forward-reference it. Contrast it with proof-goal manipulation
  (Coq/Isabelle/TLAPS) or counterexample deconstruction.
Develop an invariant
- I bet you're missing an invariant
Patch up a proof
- bugs in definitions (off by one errors)
- witnesses
- automation syntactic triggering

recursion vs. quantification story

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{Managing automation}