### 2020-09-01 osdi20-artifact/eval.tex
hey, add the paper, too

\section{Evaluation}\label{sec:eval}

Our evaluation is designed to answer two main questions:
\begin{enumerate}
  \item Do our automation-control techniques (\S\ref{sec:discipline}) and language
    improvements (\S\ref{sec:lang}) improve the developer experience? 
  \item Can our verification methodology scale to the complexity of a
    modern key-value-store data structure, and can we deliver the
    performance gains of write optimization?
\end{enumerate}

\todo{We should measure how much (and what percentage) of code we added/modified to create linear Dafny}

\subsection{Developer Experience}\label{sec:eval:effort}

\paragraph{Measuring Tedium}\label{eval:tedium}
We estimate the amount of tedium (or conversely, the efficacy of automation)
by the ratio of the lines of proof 
(e.g., lemmas, pre-/post-conditions, loop invariants)
to the lines of executable implementation code.
This is not a precise model, since it measures a completely verified artifact,
where the developers may have cleaned up temporary lines of tedium that
were typed in the course of resolving verification failures.
However, the proof text in the ``cleaned up'' code at least
reflects the tedium needed to bridge what automation
could not manage by itself.

\begin{figure}
\begin{center}
\input{data/line-counts}
\end{center}
  \vspace{-3mm}
\caption{Line counts~\cite{sloc} by major components in Figure~\ref{fig:full}.}
\label{line-counts}
\end{figure}

\begin{figure}
\begin{center}
\input{data/line-counts-micro}
\end{center}
  \vspace{-3mm}
\caption{Line counts~\cite{sloc} of two subcomponents, comparing dynamic-frame implementations with  our linear type system.\protect\footnotemark~
        Linear typing reduces the proof burden by 31-37\%}
\label{line-counts-micro}
  \vspace{-2mm}
\end{figure}

%\todo{It would be nice to have Disk, IoSystem and such like be charged to a ``framework'' category}

Figure~\ref{line-counts} gives line counts for \name, 
organized by the major components shown in Figure~\ref{fig:full}.
We see that the proof ratio for the implementation code
is 4:1, which grows to 7:1 %for the system as a whole
%(i.e., when proving crash safety).
when including all system refinement proofs (``total'').
This is comparable to the distributed, in-memory key-value store
verified in previous work~\cite{ironfleet},
which also reports a 7:1 ratio.
However, \name's implementation is 3$\times$ larger,
indicating that automated verification techniques can scale
to larger systems without super-linear effort.

The results also show that \name's implementation is 
more than 5$\times$ larger than its specification,
giving us reason to hope that the specification
is less likely to contain bugs than an unverified implementation.

Figure~\ref{line-counts-micro}\footnotetext{The implementation of the hash table (with dynamic frames and the linear type system) and search tree (with dynamic frames) coincidentally have an identical line count, despite having substantial differences.} 
compares two \name components that we wrote using
both dynamic frames and linear reasoning.
The results show that
switching to linear reasoning saves tedium,
reducing proof overhead by 31-37\%. \todo{Overhead on verification time?}
%on the hash table
%and search tree subcomponents.

\paragraph{Measuring Proof-Attempt Latency}\label{eval:discipline}
To assess the success
of our discipline for keeping automation under control (\S\ref{sec:discipline}),
we measure the time to verify individual proof
units (e.g., definitions, methods, or lemmas) 
where developers spend most of their time waiting for verification results.
This approximately measures the developer's perception of the 
latency of the verification-development cycle.

\begin{figure}
\includegraphics[width=1\columnwidth]{figures/verification-times.pdf}
  \vspace{-5mm}
\caption{Cumulative distribution of verification times of function
  definitions, implementation methods, and proof lemmas. Most
  definitions--{\twentySecFasterPctile}--verify in less than {\twentySecSlowThresholdSecs}s,
  and {\tenSecFasterPctile}--verify in less than {\tenSecSlowThresholdSecs}s.
  }
  \vspace{-4mm}
\label{verification-times}
\end{figure}

Figure~\ref{verification-times} demonstrates that
{\name} almost always exhibits interactive verification times,
with {\tenSecFasterPctile} completing in under {\tenSecSlowThresholdSecs} seconds,
and {\twentySecFasterPctile} in under {\twentySecSlowThresholdSecs} seconds.
This suggests that our timeout-averse
development policy is effective.
The figure reveals that
we did not apply our policy with perfect consistency,
as {\twentySecNumSlowVerifications} proofs do take longer than
{\twentySecSlowThresholdSecs}s to verify.
Of those, {\twentySecPctOfSlowVerifsInvolvingHeap} involve dynamic frames.

It is difficult to capture every place we applied the policy. As a proxy,
we can count places where we explicitly hid a definition.
This both overestimates the cost of the policy 
(because sometimes we hid a definition based on intuition,
before observing a timeout) and underestimates it (because automation can
be controlled with other techniques).

\begin{figure}
\includegraphics[width=1\columnwidth]{figures/automation-figure.pdf}
  \vspace{-4mm}
  \caption{{\System} manually hides {\autoOpaqueCount} definitions ({\autoOpaquePct} of all system definitions),  revealing
them {\autoRevealCount} times.
This metric reflects the effort involved in explicitly managing automation.
}
  \vspace{-3mm}
\label{automation-histogram}
\end{figure}

These approximations aside, Figure~\ref{automation-histogram}
gives an idea of the burden of controlling automation.
We hid {\autoOpaqueCount} definitions,
{\autoOpaquePct} of all definitions in the system.
These hidden definitions are manually revealed {\autoRevealCount} times.
Of the {\autoOpaqueCount}, {\autoOpaqueZeroReveals} were never revealed:
the salient properties of the definition could be exported as a postcondition
without causing timeouts.
{\autoFivePct} of hidden definitions are revealed no more than five
times; their essential features are captured in lemmas or wrapped into
higher-level definitions.
%
%\cut{The most revealed definition is IsStrictlySorted, revealed {\autoRevealMax} times.}

Developers are typically less sensitive to the latency of continuous-integration
builds that check that the system as a whole still verifies.
For \name, these take 1.8 hours of CPU time,
but thanks to the inherent parallelism of modular verification,
complete in 11 minutes on 32 cloud machines.

\subsection{Performance}\label{sec:eval:perf}

Our performance evaluation is designed for two questions:
\begin{enumerate}
  \item Does {\name} demonstrate
    the insertion-performance gains of write-optimized data
    structures?
  \item Does our linear extension produce code with
    performance comparable to hand-written code using dynamic frames?
\end{enumerate}

All experiments were run on cloud instances with directly attached physical storage.
The HDD experiments and subcomponent microbenchmarks are run on AWS EC2 d2.xlarge instances, with 4x hardware hyperthreads on a 2.4 GHz Intel Xeon E5-2676 v3.
The SSD experiments are run on AWS EC2 i2.xlarge instances, with 4x hardware hyperthreads on a 2.5 GHz Intel E5-2670 v2 and a SATA SSD.

\newcommand{\ycsbHddDataDir}{data/ycsb-hdd/golden}
\newcommand{\ycsbSsdDataDir}{data/ycsb-ssd/golden}
\newcommand{\mutablemapDataDir}{data/mutablemap/boldgolden}
\newcommand{\mutablebtreeDataDir}{data/mutablebtree/silver}

\pgfplotsset{
  GenericBarplot/.style={
    ybar=1pt,
    bar width=5.5pt,
    height = 1.75in,
    enlarge x limits=0.1,
    major x tick style = transparent,
    scaled y ticks=false,
    ylabel near ticks,
    point meta=rawy,
    nodes near coords,
    every node near coord/.append style={rotate=90, anchor=west, font=\tiny},
    nodes near coords={\pgfmathprintnumber[fixed, precision=0]{\pgfplotspointmeta}},
    area legend,
    legend columns=-1,
    label style={font=\small, align=center},
  },
  MinMidMaxFile/.style n args={3}{
    /pgfplots/error bars/y dir=both,
    /pgfplots/error bars/y explicit,
    table/y error plus expr = (\thisrow{#3} - \thisrow{#2}),
    table/y error minus expr = (\thisrow{#2} - \thisrow{#1}),
  }
}

\pgfplotsset{
  YcsbThroughput/.style={
    GenericBarplot,
    symbolic x coords = {Load, A, B, C, D, F, Cuniform},
    ylabel={Operations/Second},
    MinMidMaxFile={MinThroughput}{MedianThroughput}{MaxThroughput},
  },
  YcsbLoadThroughput/.style={
    YcsbThroughput,
    width = 1.1in,
    restrict x to domain=0:0,
    xtick = {Load},
    xlabel={(from YCSB A)},
  },
  YcsbRunThroughput/.style={
    YcsbThroughput,
    width = 3.1in,
    xmin = A,
    xmax = Cuniform,
    restrict x to domain=1:10,
    xtick = {A, B, C, D, F, Cuniform},
    xticklabels = {A, B, C, D, F, U},
    xlabel={YCSB Workload},
  },
}

\pgfplotsset{
    LinearStyle/.style={teal, fill=teal!10!white, postaction={pattern=north west lines, pattern color=teal!50!white}},
    HeapStyle/.style={violet!50!black, fill=violet!50!black!10!white, postaction={pattern=north east lines, pattern color=violet!50!black!50!white}},
}

\pgfplotsset{
    BerkeleyStyle/.style={blue, fill=blue!10!white, postaction={pattern=crosshatch dots, pattern color=blue!50!white}},
    VeriHeapStyle/.style={HeapStyle},
    VeriLinearStyle/.style={LinearStyle},
    RocksStyle/.style={red, fill=red!05!white, postaction={pattern=crosshatch, pattern color=red!50!white}},
}

\begin{figure*}
\centering
%\ref{ycsb-legend}
\begin{subfigure}[c]{\textwidth}
  \centering
  \parbox[c]{0.1\textwidth}{\subcaption{HDD}} ~
  \parbox[c]{0.7\textwidth}{
  \begin{tikzpicture}
    \begin{axis}[
        YcsbLoadThroughput,
        ymode = log,
        ymax = 100000,
        ytick = { 100, 1000, 10000, 100000, 1000000 },
        legend columns=1,
        legend to name={ycsb-legend},
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbHddDataDir/berkeleydb.csv};
      \addlegendentry{BerkeleyDB}
      \addplot+ [VeriHeapStyle] table {\ycsbHddDataDir/veribetrkv.csv};
      \addlegendentry{\name-DF}
      \addplot+ [VeriLinearStyle] table {\ycsbHddDataDir/veribetrkv-linear.csv};
      \addlegendentry{\name}
      \addplot+ [RocksStyle] table {\ycsbHddDataDir/rocksdb.csv};
      \addlegendentry{RocksDB}
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        YcsbRunThroughput,
        ymin = 0,
        ymax = 800,
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbHddDataDir/berkeleydb.csv};
      \addplot+ [VeriHeapStyle] table {\ycsbHddDataDir/veribetrkv.csv};
      \addplot+ [VeriLinearStyle] table {\ycsbHddDataDir/veribetrkv-linear.csv};
      \addplot+ [RocksStyle] table {\ycsbHddDataDir/rocksdb.csv};
    \end{axis}
  \end{tikzpicture}
  }%
  \hspace{-0.6in}%
  \ref{ycsb-legend}
  \label{subfig:ycsb-hdd}
\end{subfigure}

\begin{subfigure}{\textwidth}
  \centering
  \parbox[c]{0.1\textwidth}{\subcaption{SSD}} ~
  \parbox[c]{0.7\textwidth}{
  \begin{tikzpicture}
    \begin{axis}[
        YcsbLoadThroughput,
        ymode = log,
        ymax = 200000,
        ytick = { 100, 1000, 10000, 100000, 1000000 },
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbSsdDataDir/berkeleydb.csv};
      % \addlegendentry{BerkeleyDB}
      \addplot+ [VeriHeapStyle] table {\ycsbSsdDataDir/veribetrkv-128k.csv};
      % \addlegendentry{\name-Dynamic Frames}
      \addplot+ [VeriLinearStyle] table {\ycsbSsdDataDir/veribetrkv-linear-128k.csv};
      % \addlegendentry{\name}
      \addplot+ [RocksStyle] table {\ycsbSsdDataDir/rocksdb.csv};
      % \addlegendentry{RocksDB}
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        YcsbRunThroughput,
        ymin = 0,
        ymax = 20000,
        ytick = { 5000, 10000, 15000, 20000, 1000000 },
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbSsdDataDir/berkeleydb.csv};
      \addplot+ [VeriHeapStyle] table {\ycsbSsdDataDir/veribetrkv-128k.csv};
      \addplot+ [VeriLinearStyle] table {\ycsbSsdDataDir/veribetrkv-linear-128k.csv};
      \addplot+ [RocksStyle] table {\ycsbSsdDataDir/rocksdb.csv};
    \end{axis}
  \end{tikzpicture}
  }
  \hspace{0.9in}
  \label{subfig:ycsb-ssd}
\end{subfigure}
% %   \caption{
% %     Median throughput of YCSB workloads with 24B keys and 512B
% %     values running on a HDD and 2GB of RAM.  Load is 10M operations
% %     ($\approx$ 5GB of data) and runs are 10000 operations each.
% %     Workload ``U'' is a workload of uniformly random queries.
% %     Error
% %     bars indicate min/max of 6 runs.  Higher is
% %     better.}\label{fig:ycsb-hdd-512B}
\caption{
Median throughput of YCSB workloads values running on an HDD (\subref{subfig:ycsb-hdd}) and SSD (\subref{subfig:ycsb-ssd}) and 2GB of RAM. \name-DF is a version of our system with hand-tuned code using dynamic frames reasoning.  Load is 10M operations
($\approx$ 5GB of data) and runs are 10000 operations each.
Error bars indicate min/max of 6 runs.  Higher is better.
On an HDD (\subref{subfig:ycsb-hdd}), \name insertions are over $25\times$ faster than in
BerkeleyDB, but lag RocksDB, by about $9\times$.
\name queries are are about $4\times$ slower than both.
On an SSD (\subref{subfig:ycsb-ssd}), \name still beats BerkleyDB on random
insertions, but \name queries are slower than BerkeleyDB.
}
\label{fig:ycsb}
\end{figure*}

\subsubsection{YCSB}

\Cref{fig:ycsb} show throughput for \name,
BerkeleyDB, and RocksDB on the YCSB benchmarks~\cite{ycsb} on hard drive and SSD,
including the load phase for workload A, and a uniformly random query
workload (labeled as workload ``U'').

There are three main take-aways from these measurements.  First, \name
demonstrates the performance gains of using a write-optimized data
structure.  For the load phase on hard drive, which consists of a pure
random insertion workload, \name is over $25\times$ faster than
BerkeleyDB.  Even on SSD, where random I/O is much cheaper, \name
modestly outperforms BerkeleyDB on insertions.

In contrast, \name is roughly $8\times$ slower than RocksDB on hard
disk and $4\times$ slower on SSD.  We think this gap is due to two main
factors.  First, our implementation performs a sync whenever its
on-disk log fills up, whereas RocksDB performs sync only when the
application requests them.  Second, due to memory fragmentation and
other overheads, our effective cache size is approximately $4\times$
smaller than Rocks.  Ths is because Rocks primarily uses the OS page
cache, which can pack 4KB pages in memory, whereas we use an internal
(and, hence, verified!) cache of nodes.

The second main take-away is that queries in \name are a constant
factor slower than in a \btree: roughly $4\times$ slower than
BerkeleyDB on both hard drive and SSD.  It is well known that
write-optimized data structures need additional optimizations to match
\btree point-query performance.  For example, RocksDB is
also slower than BerkeleyDB on SSD and has relatively mixed
performance on hard disk.  The primary bottleneck for \name query
performance is that each cache miss loads an entire node, and nodes
are large (2MB on hard disk and 128K on SSD) in order to amortize I/O
overheads.  BerkeleyDB nodes, on the other hand, are only 16KB.
Industrial-grade \beptrees solve this problem with
``sub-nodes'', but we have not yet implemented this optimization.

The final take-away is that at a macro-level our linear implementation has essentially
the same performance as the version with hand-tuned code
using dynamic frames reasoning.  
%This demonstrates that our linear type system extension to
%Dafny enables us to write algorithms and data structures as if we are
%using immutable variables, but compile down to efficient,
%update-in-place code.

Overall, we conclude that \name demonstrates that a verified system
can achieve the performance gains of a write-optimized storage system,
but it needs further optimization to match highly-tuned commercial
implementations.

\begin{figure}
\centering
\ref{micro-legend}
\begin{subfigure}[t]{.25\textwidth}
  \begin{tikzpicture}
    \begin{axis}[
        GenericBarplot,
        enlarge x limits=0.2,
        width = 1.8in,
        symbolic x coords = {insert, readpositive, readnegative, remove},
        xtick = {insert, readpositive, readnegative, remove},
        xticklabels = {Insert, Pos. Query, Neg. Query, Delete},
        x tick label style={rotate=90},
        ymin = 0,
        ymax = 8300000,
        ytick = {0, 2000000, 4000000, 6000000, 8000000},
        yticklabels={0, 2M, 4M, 6M, 8M},
        ylabel={Operations/Second},
        legend to name={micro-legend},
        MinMidMaxFile={min}{med}{max},
      ]
      \addplot+ [HeapStyle] table {\mutablemapDataDir/master.csv};
      \addlegendentry{Dynamic Frames}
      \addplot+ [LinearStyle] table {\mutablemapDataDir/linear.csv};
      \addlegendentry{Linear}
    \end{axis}
  \end{tikzpicture}
  \subcaption{Hash table}
        \label{subfig:micro-hashtable}
\end{subfigure}~%
\begin{subfigure}[t]{.2\textwidth}
  \centering
  \parbox{1.7\textwidth}{
    \vspace{-2.475in}
  \begin{tikzpicture}
    \begin{axis}[
        GenericBarplot,
        enlarge x limits=0.5,
        width = 1.3in,
        symbolic x coords = {write,read},
        xtick = {write, read},
        xticklabels = {Insert, Query},
        x tick label style={rotate=90},
        ymin = 0,
        ymax = 300000,
        ylabel={Operations/Second},
        MinMidMaxFile={min}{med}{max},
      ]
      \addplot+ [HeapStyle] table {\mutablebtreeDataDir/8000000-master.csv};
      \addplot+ [LinearStyle] table {\mutablebtreeDataDir/8000000-linear.csv};
    \end{axis}
  \end{tikzpicture}}\\
        \subcaption{Search tree}
        \label{subfig:micro-btree}
\end{subfigure}
        \caption{Median throughput of subcomponent microbenchmarks.
Higher is better. Error bars are min/max of 6 runs.
%Our linear hash table (\subref{subfig:micro-hashtable}) is slightly faster,
%or at least comparable, than our hand-optimized implementation using
%dynamic frames reasoning. Throughput of successively inserting
%64 million key-value pairs, performing 64 million positive and
%negative queries, and emptying the hash table by removing each
%key. 
%Our linear in-memory search tree (\subref{subfig:micro-btree}) is comparable with our
%hand-optimized update-in-place version using dynamic frames reasoning.
%Throughput of inserting 8 million key-value pairs, and performing 8 million negative queries.
}\label{fig:datastructures-micro}
\end{figure}

\subsubsection{Linear Data Structures}

\Cref{fig:datastructures-micro} shows the performance of
our linear and dynamic-frames-based hash-table and search-tree implementations.
%
The main take-away from both experiments is that, even in
microbenchmarks, the linear and non-linear implementations have very
comparable performance.

The hash table benchmark inserts 64 million key-value
pairs, then performs as many positive and negative queries, and then
deletes everything in the hash table.  The linear version is slightly
faster than the non-linear version, except for deletes, which are
slightly slower.  We suspect the speedup comes primarily from the lack
of shared pointer overhead.

The search-tree benchmark measures the time to insert 8 million key-value
pairs and then query them all.  Performance for the linear version is
close to the non-linear version, but slightly faster for queries and
slightly slower for inserts.  We believe queries are faster due to the
elimination of shared pointer overheads, and the insertions are slower
due to the overheads of destructing nodes on the way down the tree and
reconstructing them on the way back up.  
%We could eliminate this
%overhead by adding support for mutably borrowing references to our
%linear type system.  The non-linear version was also written to be
%tail recursive.  This is not possible in the linear version, since it
%needs to reconstruct nodes as it unwinds the recursion.  Mutable
%borrow support would also eliminate this overhead.

Overall, we conclude that our linear type system enables us to
construct performant code without the challenges of dynamic frame
reasoning.

%% \begin{figure}
%% \includegraphics[width=1\columnwidth]{figures/btree-perf.pdf}
%% \caption{linear/repr perf}
%% \label{btree-perf}
%% \end{figure}